\section{Cristina and centers}
Cristina is interested in a metric space (X, d), where all distances d(x, y) are
either 0 (in which case x = y), 1 or 3. Further, all distances are symmetric and obey the triangle inequality, that is $d(a, b) + d(b, c) \geq d(a, c)$. Cristina wants to cluster the points of X. Unfortunately, she does not yet know how many clusters she has to use. She therefore will find a permutation $\Pi(X)$ and use, for every $k \in \{1, ..., |X|\}$, the first k elements $C =
{\pi(X)_1, ..., \pi(X)_k}$ of the permutation as centers.
\paragraph*{Goal:} For any such k, the output should be optimal with respect to following objective function:
\begin{equation}
\max_{x \in X} \min_{c \in C} d(x,c).
\end{equation}
\paragraph*{Hint:} Use a greedy algorithm. The running time should be no larger than $O(|X|^2)$

\subsection{The algorithm}
This problem can be seen as a variant of the well-known \cite{K-center} K-centers decision problem. I have designed a greedy algorithm that solves the above problem.
It creates (and returns to Cristina) the desired permutation in this way.
\begin{enumerate}
	\item $C:= [\,]$
	\item pick a random point c\_first and add it to the list of centers C.
	\item for every point v $\in$ V (v $\notin C$), compute the minimum distance from C to v 
	\item pick the point c\_next with highest distance from the points in C
	\item add c\_next to C, "relax" the distances of points outside C, and continue the process from (3) until $|V| = |C|$
\end{enumerate}

\subsubsection{Source-code}
I have implemented the algorithm in Python 3.7 and the source code is attached as ex1.py. Moreover, I provide the pseudo-code as ex1.txt.

\subsubsection{Running time}
The algorithm adds $|X|$ points to C. How much costs adding a point to C? A point is added to C if, among all the points v $\notin$ C, it has the maximum distance from the ones in C. At each iteration we update the minimum distance from all points to C, simply considering if the new center can "decrease" the current distance of the point from C. It costs $|X - 1|$ for the second center, $|X - 2|$ for the third and so on. Since the iterations are $|X|$ and each one costs $O(|X|)$ we have the cost is $O(|X^2|)$.
\subsubsection{Proof of correctness}
At each iteration the algorithm updates (if possible) the current maximum distance from the list of centers. This leads to a non increasing sequence of max\_values = $\{d_1, d_2, ..., d_{|X|}\}$. Given an optimal solution O, we have to prove that the sequence computed with the greedy algorithm above is such that for every $i \in \{1, 2, ..., |X|\}$, $d_g[i] = d_o[i] $, where $d_g[i]$ is the maximum distance from C in the greedy algorithm and $d_o[i]$ is the maximum distance from C in the optimal solution at iteration i. Since my greedy algorithm guarantees \cite{K-center} that for every k, $d_g[k] \leq 2 \cdot d_o[k]$ and since the only possible values of d are 0, 1 or 3, we have these possible cases:
\begin{itemize}
	\item $d_o[k] = 0 \Rightarrow  0 \leq d_g[k] \leq 2\cdot0 = 0 \Rightarrow d_g[k] = 0$ OK
	\item $d_o[k] = 1 \Rightarrow  1 \leq d_g[k] \leq 2\cdot1 = 2 \Rightarrow d_g[k] = 1 $, OK since 1 is the only value that d can assume s.t. $1 \leq d \leq 2$
	\item $d_o[k] = 3 \Rightarrow  3 \leq d_g[k] \leq 2\cdot3 = 6 \Rightarrow d_g[k] = 3 $, OK since 3 is the only value that d can assume s.t. $3 \leq d \leq 6$
\end{itemize}
And this is true for every k.